\documentclass[11pt]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{amsmath, amssymb, bm}
\usepackage{hyperref}

\title{Real ESDF-Time Replacement for the Static \texttt{map\_idx} Shortcut}
\author{REACT stage-3.7+ candidate fix}
\date{2026-05-22}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\esdf}{\mathcal{D}}

\begin{document}
\maketitle

\section{The current shortcut and its noise floor}

\texttt{DynamicYOPOWrapper.\_\_getitem\_\_} (\texttt{policy/yopo\_dataset.py:385})
assigns a random static \texttt{map\_idx}:
\begin{verbatim}
map_idx = int(np.random.randint(0, self.env_num_static))
\end{verbatim}

The trainer then evaluates \texttt{YOPOLoss.safety\_loss} against the ESDF of
\emph{that randomly selected} static map. The actual dynamic-bake's static
point cloud --- different \texttt{env\_X} folder, different forest layout ---
is silently ignored.

\subsection{Why this matters}

For a sample with predicted endstate $\bm{p}_i^{\text{pred}}$ and the (random)
queried ESDF $\esdf_{m^*}$ where $m^* \sim \mathrm{Uniform}(0, E_{\text{static}})$,
the supervision is
\begin{equation}\label{eq:safety-current}
\mathcal{L}_{\text{safety}}^{\text{cur}}(\bm{p}_i^{\text{pred}})
\;=\;
\softplus\bigl(d_{\text{safe}}^{\text{stat}} - \esdf_{m^*}(\bm{p}_i^{\text{pred}})\bigr).
\end{equation}
But the \emph{true} static obstacles in this sample live at positions
$\{\bm{q}_j\}_{j}$ from the dynamic-bake's own pointcloud (a different forest).
We are penalising the network for ``collisions'' that don't exist in this
scene, and ignoring real collisions that do.

\subsection{Decomposing the noise}

Write the desired ``true'' supervision as
$\mathcal{L}_{\text{safety}}^{\text{true}} = \softplus(d - \esdf_{\text{true}}(\bm{p}_i^{\text{pred}}))$.
The expected gradient mismatch between current and true at a query point $\bm{p}$ is
\begin{equation}\label{eq:mismatch}
\bm{\eta}(\bm{p})
\;=\;
\mathbb{E}_{m^*}\!\Bigl[
   \nabla_{\bm{p}} \mathcal{L}_{\text{safety}}^{\text{cur}}(\bm{p}) -
   \nabla_{\bm{p}} \mathcal{L}_{\text{safety}}^{\text{true}}(\bm{p})
\Bigr].
\end{equation}
If the static-forest distribution is sufficiently homogeneous (i.i.d.\ tree
positions, same density), $\Norm{\mathbb{E}_{m^*}[\nabla \esdf_{m^*}]}$
$\to 0$ as $E_{\text{static}} \to \infty$, so $\bm{\eta}$ has \emph{zero
expected bias} --- but \emph{non-zero variance}. The variance shows up as
gradient noise on every dynamic-batch step.

For our v3 forests with $\rho_{\text{tree}}\!\approx\!0.05$ trees per m$^2$
and $r_{\text{tree}}\!\approx\!0.3\,$m, the per-sample $\Norm{\nabla \esdf}$
standard deviation is on the order of $\Norm{\nabla \esdf}_{\text{rms}} \approx 1$
in regions away from the tree boundary, dropping to $\esdf$-sign-flip
indicator near boundaries. The noise contribution to
\texttt{stat\_traj\_tail} on dynamic batches is bounded but real.

\section{Direct fix: bake the dynamic scene's own ESDF}

Each \texttt{dataset\_dynamic/v$X$/env\_$E$/pointcloud.ply} already lives on
disk. Computing an ESDF voxel grid from it is a standard one-time precomputation
(use the same KdTree machinery \texttt{Simulator/}'s \texttt{maps.hpp} already
employs). For $60{\times}60{\times}15$\,m at $0.1$\,m resolution this is
$\sim$5.4\,M voxels per env; 40 envs $\to$ $\sim$216\,M voxels $\to$
$\sim$1\,GB at float32. Acceptable.

\subsection{Time-aware ESDF over the moving spheres}

For full ``ESDF-time'' supervision (matching the formulation hinted at in
\texttt{02\_multi\_waypoint\_extension.tex}), define
\begin{equation}\label{eq:esdf-time}
\esdf_{\text{full}}(\bm{p}, t)
\;=\;
\min\Bigl\{
   \esdf_{\text{static}}(\bm{p}),
   \;
   \min_{m=1,\dots,M}
   \bigl[\Norm{\bm{p} - \bm{p}_m^{\text{obs}}(t)} - r_m\bigr]
\Bigr\},
\end{equation}
where $\bm{p}_m^{\text{obs}}(t) = \bm{p}_m^{\text{obs}}(0) + t\bm{v}_m^{\text{obs}}$
(linear motion --- matches the bake). $\esdf_{\text{static}}$ is the precomputed
static-cloud distance field. The full safety supervision becomes
\begin{equation}\label{eq:safety-true}
\mathcal{L}_{\text{safety}}^{\text{full}}(\bm{p}_i, t_i)
\;=\;
\softplus\bigl(d_{\text{safe}}^{\text{stat}} - \esdf_{\text{full}}(\bm{p}_i, t_i)\bigr).
\end{equation}

The gradient is well-defined wherever the inner $\min$ is unique --- in
practice we use a smooth-min approximation
$\widetilde{\min}_\beta(x_1, \dots, x_k) = -\beta^{-1}\log\sum_j e^{-\beta x_j}$
with $\beta \sim 10$ to stabilise near the ``static vs. dynamic surface''
boundary.

\subsection{Gradient through the time argument}

The trajectory's $t^{(n)}$ are differentiable functions of the network output
(the time-head in \texttt{gru\_decoder.py} predicts them). Differentiating
\eqref{eq:esdf-time} wrt $t$:
\begin{equation}
\frac{\partial \esdf_{\text{full}}}{\partial t}\Big|_{m^*}
\;=\;
\hat{\bm{u}}_{i,m^*} \cdot \bm{v}_{m^*}^{\text{obs}}
\quad
\text{(only at the dynamic minimum; static branch is }t\text{-independent)}.
\end{equation}
This gives the network a direct signal that ``shifting the waypoint timing
moves you closer to or farther from the moving obstacle'', which neither the
current \texttt{safety\_loss} nor the current \texttt{motion\_reshaped} alone
provide.

\section{Implementation cost}

\begin{itemize}
  \item \textbf{Bake-time:} add $\esdf_{\text{static}}$ precomputation as a
        post-bake step in \texttt{Simulator/src/src/dataset\_generator.cpp};
        ship as \texttt{esdf.bin} (binary float grid) alongside
        \texttt{pointcloud.ply} per env. One-off cost; v3 already has the
        pointcloud.
  \item \textbf{Training-time:} new \texttt{policy/utils/esdf\_query.py} doing
        trilinear interp on the float grid. CUDA kernel optional; for v3 we
        can start with CPU since the env count is moderate.
  \item \textbf{Loss replacement:} swap \texttt{YOPOLoss.safety\_loss}'s
        \texttt{ESDFQuery} (which currently reads the \emph{static-dataset}
        ESDF indexed by \texttt{map\_idx}) for the dynamic-scene-correct one.
        Strict regression: when training on the static dataset
        (\texttt{dynamic\_ratio=0}), the static-ESDF path is byte-identical
        to today.
\end{itemize}

Estimated work: 3--4 days; this is the cheapest of the three supervision
upgrades (compared to multi-waypoint head and camera-aware bake) but its
expected impact is the smallest --- it fixes a gradient \emph{noise} source,
not a structural saturation. Recommend doing it \emph{after} multi-waypoint to
clean up the residual once Option B is in place.

\section{Expected impact (back-of-envelope)}

In the current setup the static-safety mismatch contributes a $\sim$5--10\%
share of total dynamic-batch gradient norm. Eliminating that noise should
reduce \texttt{stat\_traj\_tail} on dynamic batches by a similar fraction
($\sim$$0.07$ absolute reduction from $\sim$$3.76$ to $\sim$$3.69$). It is
\emph{unlikely} to move \texttt{dyn\_dyn} more than 1--2\% because the
saturation argument in \texttt{01\_collision\_loss\_saturation.tex} is
independent of this fix.

\end{document}
